---
title: 'dyngame: an R package to solve discounted  dynamic games with transfers'
output:
  html_document:
    highlight: tango
    keep_md: yes
    theme: readable
    toc: yes
---

The R package `dyngame` can be used to solve discounted stochastic games with perfect monitoring in which monetary transfers can be used. The package computes the payoff set of all pure strategy subgame perfect equilibria and characterizes Pareto-optimal SPE. It implements the  policy elimination algorithm described in the paper "Discounted Stochastic Games with Voluntary Transfers" by Susanne Goldluecke and Sebastian Kranz.


The code below shows how to install the required R packages and contains the dynamic Cournot game example from the paper. Look in the example section in the packages github folder:

https://github.com/skranz/dyngame/tree/master/examples

for the source code files of the example.

Unfortunatley, the package documentation is still not well developed.

# Install packages

Run the following code in your R console to install the required packages from the web.

```{r eval=FALSE}
if (!require(devtools))
  install.packages("devtools")

devtools::install_github("skranz/skUtils")
devtools::install_github("skranz/rgmpl")
devtools::install_github("skranz/dyngame")




```


# Cournot example with stochastic water reservers from the paper


## Game definition
```{r include=FALSE}
library(dyngame)
```

```{r "game", tidy=TRUE}
library(dyngame)

# Some constants used in the game specification

# type of reserve increment
DET = 1  # deterministic reserve increment
UNIF = 2 # uniformely distributed reserve increment

# type of solution
COLL = 1 # collusive solution (optimal dynamic game equilibrium)
MON = 2  # integrated monopoly solution


# a quicker translation function from states vectors to state indices
xv.to.x = function(m,xvm) {
  return( (xvm[,1])*m$xv.dim[[2]]+xvm[,2]+1)
}

# Cournot duopoly with stochastic water reserves. Example from paper
cournot.reserves.game = function(
           x.cap=20, K=20, x.inc=2, delta=2/3,inc.min=0,inc.max=2,
           sol.type=COLL,inc.type=DET,para=NULL) 
{
  
  # If parameters are given as a list just copy them all
  # into the local environment
  if (!is.null(para)) {
    copy.into.env(source=para)
  }
    
  # act.fun specifies the action set for a given state xv
  # val = numerical values of actions
  # lab = labels of actions
  act.fun = function(xv,m=NULL) {
    restore.point("act.fun")
  	val = list(0:xv[1],0:xv[2])
		lab = val
		
    list(val=val,lab=lab, i=c(1,2))
  }
  
  # States can be grouped into sets of states with same
  # action sets. This will speed up computations.
  # In our game each state has a different action set
  x.group = function(xvm,m=NULL) {
    # return a different group index for each state
    x.group = 1:NROW(xvm)
    x.group
  } 
	
  # Stage game payoff function
  # Specifies the stage game payoffs as a function
  # of the action profile av and state xv
  # The function must be vectorized, i.e. avm and xvm
  # are matrices of action profiles and states
  g.fun = function(avm,xvm,m=NULL) {
    restore.point("g.fun");
  	
		rownames(avm)=rownames(xvm)=NULL
  	
    # compute outputs and prices
		q = avm
  	P = K-q[,1]-q[,2]
  	P[P<0]=0
    # return profits
  	pi = P*q 
  	pi
  }
  
  # Some additional statistics of the solution that we might be interested in.
  # Here we want to know about price, total output, joint profits,
  # consumer surplus, and total welfare 
  extra.sol.fun = function(avm,xvm,m=NULL) {
    q   = avm
    P   = K-q[,1]-q[,2]
    P[P<0]=0
    pi = P*q 
    Q  = q[,1]+q[,2]
    CS = (K-P)*pmin(Q,K)*(1/2)
    Pi = pi[,1]+pi[,2]
    W  = CS+Pi
    
      
    mat = cbind(q,P,Q,CS,Pi,W)
    colnames(mat)[1:2]=c("q1","q2")
    return(mat)
  }
    
  # We do not consider a multistage game
  x.stage.fun = NULL
    
  # State transitions
  # For a matrix of action profiles and states specifies
  # the matrix of state transitions
  tau.fun = function(avm,xvm,m=NULL) {
  	#restore.point("tau.fun")

		rownames(avm)=rownames(xvm)=NULL
  	tau = matrix(0,NROW(avm),m$nx)  	  	
 
		# Firms reserves are restored by a fixed amount x.inc
		if (inc.type==DET) {
      
      # matrix of resulting state values
			new.xvm = with.floor.and.ceiling(xvm - avm + x.inc ,floor=0,ceiling=x.cap)

			# Translate state values to state indices
      new.x = xv.to.x(m,new.xvm)
      
      # Set transition probabilities to 1 for resulting states 
			ind.mat = cbind(1:NROW(xvm),new.x)
			tau[ind.mat]=1
			
		# Alternatively, firms reserves are restored by a uniformely 
    # distributed integer amount between 0 and x.inc
		} else if (inc.type==UNIF) {
      
      # loop through all combinations of independently
      # distributed restore amount draws
			for (x.add1 in inc.min:inc.max) {
				for (x.add2 in inc.min:inc.max) {
					x.add = c(x.add1,x.add2)				
					xvm1 = with.floor.and.ceiling(xvm[,1] - avm[,1] + x.add1 ,floor=0,ceiling=x.cap)
					xvm2 = with.floor.and.ceiling(xvm[,2] - avm[,2] + x.add2 ,floor=0,ceiling=x.cap)
				
					# Translate state values to state indices
          new.x = xv.to.x(m,cbind(xvm1,xvm2))
          
          # Add probability of restore draw to transition matrix
					ind.mat = cbind(1:NROW(xvm),new.x)
					tau[ind.mat]=tau[ind.mat] +1/((inc.max-inc.min+1)^2)
				}
			}	
		}
  	
   	tau
  }
  
  integrated = sol.type == MON
  
  # return required information for the dynamic game
  list(n=2,
       delta=delta,
       integrated = integrated,
       xv.val = list(0:x.cap,0:x.cap),
       act.fun=act.fun,
       g.fun=g.fun,
       tau.fun=tau.fun,
       x.stage.fun = x.stage.fun,
       x.group=x.group,
       extra.sol.fun=extra.sol.fun)
}

```


## Initializing and solving game

We now solve the game using the parameterizations in our paper.
```{r tidy=TRUE, cache=TRUE}

NO.LABELS = FALSE

# Turn off restore points to speed up computation
set.storing(FALSE)

# Specify parameters 
para = list(
  delta=2/3,
  K=20,
  x.cap=20,
  x.inc=3,
  inc.min=3,
  inc.max=4,
  inc.type=UNIF,
  sol.type=COLL
)

# init game
my.game = cournot.reserves.game(para=para)
mc = init.game(my.game = my.game)

# solve for collusive solution
mc = solve.game(mc)

# Solution table with one row per state
sol = print.sol(mc)
head(sol)
```



## Analyse solution graphically

The code below performs the graphical analysis
```{r}
# Analyse solution graphically
par(mar=c(5,5,2,2))

# Equilibrium prices
state.levelplot(mc,z=mc$extra.sol.cur[,"P"],
                arrows=FALSE,cuts=10,xrange=c(0,20),zlim=c(7,20),
                main="Prices under Collusion",
                xlab="Reserves firm 1", ylab="Reserves firm 2",
                reverse.colors=TRUE)

state.levelplot(mc,z=mc$extra.sol.cur[,"P"],
                arrows=FALSE,cuts=10,xrange=c(0,20),zlim=c(8,20),
                main="Prices under Collusion",
                xlab="Reserves firm 1", ylab="Reserves firm 2",
                reverse.colors=TRUE,col.scheme = "grey")

# Punishment payoffs

V = mc$sol.mat[,"v1"]+mc$sol.mat[,"v2"]
state.levelplot(mc,z=V,arrows=FALSE,cuts=10,xrange=c(0,20),
                main="Sum of punishment payoffs",
                xlab="Reserves firm 1", ylab="Reserves firm 2",
                reverse.colors=!TRUE,col.scheme = "grey")
                
V = mc$sol.mat[,"v1"]+mc$sol.mat[,"v2"]
state.levelplot(mc,z=mc$sol.mat[,"v1"],arrows=FALSE,cuts=10,xrange=c(0,20),
                main="Punishment payoffs player 1",
                xlab="Oil reserves firm 1", ylab="Oil reserves firm 2",
                reverse.colors=!TRUE,col.scheme = "grey")

```

## Monopoly solution

  # Monopoly solution
  mon = clone(mc)
  mon$integrated = TRUE
  mon = solve.game(mon,delta=delta)
  
  par(mar=c(5,5,2,0))
  state.levelplot(mon,z=mon$extra.sol.cur[,"P"],arrows=FALSE,cuts=10,xrange=c(0,20),
                  main="Prices under Monopoly",zlim=c(8,20),
                  xlab="Oil reserves firm 1", ylab="Oil reserves firm 2",
                  reverse.colors=TRUE)
                  
   
 }
  
